---
title: "Visualizing Abortions in Denmark"
author: "Emil Buus Thomsen and Rasmus Vesti Hansen"
date: "02/01/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



```{r Loading packeges}

library(tidyverse)
library(gganimate) # This is downloadid in order to make visualizations
library(shiny) #This is downloaded in order to make interactive visualizations
library(plotly) #This is also needed for Visualizations
library(gifski) #This is also needed for Visualizations

```




Now that we have the tools we need from the library, we will download the data we would like to visualize

```{r uploading files}
finaldflink <- "https://raw.githubusercontent.com/Digital-Methods-HASS/au557916_Thomsen_Emil/main/final_df.csv"

final_df <- read_csv(url(finaldflink))
#Loading our final_df from our github repository 

#head(final_df,26)
#Checking if the data set looks as it should
```





The code chunk below uses our final dataset to show the development in abortions per capita in procent along with the average income in the different municipalities in Denmark.

```{r Visualization 1. Devellopment if income and abortions 1995-2018}
options(scipen=999) #By adding this line of code we disable the scientific way of numbers, so it is easier to read.
ggplot(final_df, aes(x = av_Income, y = allAges, size = av_Income))+ #creating the the plot and deciding what the axis should represent. And setting size of the dots to to be av income.
  geom_point() + #Making the points dots in the visualization
  transition_time(Year) + # The vizualization should change acording to the years
  labs(title = "Visualization 1: Income and Abortions in {closest_state}", x = "Average Income", y = "Abortions per capita in %") + #Giving labels to the visualization.
  transition_states(Year, transition_length = 12, state_length = 1, wrap = TRUE) +
   theme(axis.text.x = element_text(size = 12), #here we change the size of the labels.
          axis.text.y = element_text(size = 12),
          text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))

```


The three next visualizations are made with shinyApp. The inspiration and help to use it comes from stackoverflow.com:
https://stackoverflow.com/questions/38917101/how-do-i-show-the-y-value-on-tooltip-while-hover-in-ggplot2 (2017)

This visualization shows the growth in income in the differen municipalities from your 1995 to 2018.
```{r Visualization 2. Growth in income per in each municipality}
ui <- fluidPage( #fluid page creates a layout which is fluid.
    plotlyOutput("distPlot")
)


server <- function(input, output) {
   output$distPlot <- renderPlotly({
      qplot(x = av_Income, y = Year, size = av_Income, color = Municipalities, data = final_df) +  #Here we use a quick plot
       labs(title = "Visualization 2: Income growth", x = "Average Income in Municipalities", y = "Years")+ #making titles
      geom_point() #Sets the points to dots.
   })
}

shinyApp(ui = ui, server = server) #We run our shiny app.

```



This visualization only shows the municipalities of Lolland, Langeland, Gentofte and Rudersdals data for the years 1995 to 2018 and their development in abortions per capita as these Municipalities are approximately the richest 2 municipalities and the poorest 2
```{r Visualization 3. Devellopment for richest and poorest municipalities}

small_big_df <- final_df[(final_df$Municipalities == "Lolland" | final_df$Municipalities =="Gentofte" | final_df$Municipalities =="Rudersdal" | final_df$Municipalities == "Langeland"),] #creating a data frame that only consist of 4 municipalities.

#small_big_df


ui <- fluidPage(
    plotlyOutput("distPlot")
)

server <- function(input, output) {
   output$distPlot <- renderPlotly({
      qplot(x = Year, y = allAges, size = av_Income, color = Municipalities, data = small_big_df) + 
       labs(title = "Visualization 3: Poor and Rich", x = "Years", y = "Abotions per capita in %")+
      geom_point()
   })
}

shinyApp(ui = ui, server = server)




```




This code chunk below a static interactive plot to investigate the development in abortions per capita for every year in each municipality.

```{r Visualization 4.  scatter plot, abortions per person in all years}

#Scatter plot that shows abortions per person in all years and for all Municipalities

ui <- fluidPage(
    plotlyOutput("distPlot")
)

server <- function(input, output) {
   output$distPlot <- renderPlotly({
      qplot(allAges, Year, size = av_Income, color = Municipalities, data = final_df) + 
      labs(title = "Visualization 4: Abortions per capita from 1995-2018", x = "Abortions per capita in %", y = "Years") +
      geom_point()
   })
}

shinyApp(ui = ui, server = server)
```

