---
title: "HW5 - Make Data Move"
author: ""
date: "05/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Explore the recent global developments with R
Today, you will load a filtered gapminder dataset -  with a subset of data on global development from 1952 - 2007 in increments of 5 years - to capture the period between the Second World War and the Global Financial Crisis. 

**Your task: Explore the data and visualise it in both static and animated ways, providing answers and solutions to 7 questions/tasks below.**

## Get the necessary packages
First, start with installing the relevant packages 'tidyverse', 'gganimate', and 'gapminder'.

```{r libraries, echo = FALSE}
library(tidyverse)
library(gganimate) # 
library(gapminder)
library(gifski) #Loading packeges for visualizations
```

## Look at the data
First, see which specific years are actually represented in the dataset and what variables are being recorded for each country. Note that when you run the cell below, Rmarkdown will give you two results - one for each line - that you can flip between.
```{r}
unique(gapminder$year) #Here I look at the unique times years is displayed in the gapminder dataset
head(gapminder) #Looking at the start of the gapminder dataset
tail(gapminder) #Looking at the end of the gapminder dataset

```
The dataset contains information on each country in the sampled year, its continent, life expectancy, population, and GDP per capital.

Let's plot all the countries in 1952.
```{r 1952}
theme_set(theme_bw())  # set theme to white background for better visibility

ggplot(subset(gapminder, year == 1952), aes(gdpPercap, lifeExp, size = pop)) + #I use the ggplot to make visualizations. I am using "subset to look at the year year 1952 in the gapminder data set
  # aes is used to make the astetics. we set gdpPercap as the x-axis, lifeExp the y-axis and pop as the size
  geom_point() + #geom_point() gør at vi får scatter plots altså prikker i vores kordinatsystem
  scale_x_log10() #convert x to log scale
```
We see an interesting spread with an outlier to the right. Answer the following questions, please:

Q1. Why does it make sense to have a log10 scale on x axis?

If we do not ad the "log10()" line the we will still get a graph, but since there is one value that lies a long way away from the others, we will get a very long graph, which is not easy to read.
Instead if we scale the x-axis we make the graph smaller, and also easier to comprehend as the different values are more distinguishable from each other.



Q2. What country is the richest in 1952 (far right on x axis)? 

To find out which country is the richest in 1952, we must use the "gdpPercap" column to read which is the largest har the largest number and is there by the richest country.


```{r, Q2}
gapminder %>%  #We choose to look at the gapminder dataset and add pipes to it, so we can send an output of one fucntion directly to another.
  filter(year == 1952) %>%  #First we filter the years in tha dataset so vi are only looking at data from 1952
  select(country , gdpPercap) %>% #then we select or choose to look at the countries and their gdpPercap, as everything else is irelevant right now
  arrange(desc(gdpPercap)) #lastly, we arange our data in a decending order so we have the richest nation first.

```
From the above we can read that the richest country by far in 1552 was Kuwait.



You can generate a similar plot for 2007 and compare the differences
```{r 2007}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() 
```
The black bubbles are a bit hard to read, the comparison would be easier with a bit more visual differentiation.



Q3. Can you differentiate the continents by color and fix the axis labels?

```{r, Q3}
ggplot(subset(gapminder, year == 2007), aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) + #here i use the same line of code as above and ad "color = continent" to color the different continents.
  #further more we add x and y, so it is easier to change lator in the code.
  geom_point(alpha = 0.5) + # alpha gives us the opportunity to make the points more transparent on a scale from 0 -1
  labs(title = "Coutries in 2007", x = "GDP Per Capita", y = "Life Expectancy") + #Here we use labs to give the visualization a title and rename the x and y axis.
  scale_x_log10() + # convert x to log scale
  scale_size(range = c(1, 15), name="Population") #Here we scale up the size of the points in the visualization so they are easier to see. Further more we are renaming pop to Population

```

Q4. What are the five richest countries in the world in 2007?

```{r, Q4}
gapminder %>%  # We look at gapminder and use the pipes (%>%) as arguments for the next lines
  filter(year == 2007) %>%  #We filter on the year 2007
  select(country , gdpPercap) %>% #and chose to look at country and gdpPercap
  arrange(desc(gdpPercap)) %>%  #aranging the data in decending order
  head(5) #and looking at the first 5.

```
By using the almost the same code as in Question 2 and adding head 5 we can show the 5 richest countries in 2007
by reading from the chart we can see that the richest contries are Norway, Kuwait, Singapore, Unitid States and Ireland.


##  Make it move!

The comparison would be easier if we had the two graphs together, animated. We have a lovely tool in R to do this: the `gganimate` package. And there are two ways of animating the gapminder ggplot.

### Option 1: Animate using transition_states() 

The first step is to create the object-to-be-animated
```{r anim1 start}
anim <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) + #Here we make an object named anim, where we use the illustration tool ggplot and use it on the gapminder dataset. The aestetics of the illustration is afterwards determined with gdppercap as the x-axis. lifeExp as the y-axis and population as size of the points.
  geom_point() + #here we choose the style of the plot to be geom.
  scale_x_log10()  # convert x to log scale
anim #call the object.
```

This plot collates all the points across time. The next step is to split it into years and animate it. This may take some time, depending on the processing power of your computer (and other things you are asking it to do). Beware that the animation might appear in the 'Viewer' pane, not in this rmd preview. You need to knit the document to get the viz inside an html file.

```{r anim1}
anim + transition_states(year, 
                      transition_length = 1,
                      state_length = 1)#+
 #   labs(title = "{closest_state}") +
 # transition_states(year, transition_length = 12, state_length = 1, wrap = TRUE)

#The last two lines, have I added and commented out, but they fix the problem, makes the illustration, more fluid and puts changing headlines above the illustration - more explanation futher down.

```
Notice how the animation moves jerkily, 'jumping' from one year to the next 12 times in total. This is a bit clunky, which is why it's good we have another option. 


### Option 2 Animate using transition_time()
This option smoothes the transition between different 'frames', because it interpolates and adds transitional years where there are gaps in the timeseries data.

```{r anim2}
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() + # convert x to log scale
  transition_time(year)
anim2
```

The much smoother movement in Option 2 will be much more noticeable if you add a title to the chart, that will page through the years corresponding to each frame.


Q5 Can you add a title to one or both of the animations above that will change 
 in sync with the animation? [hint: search labeling for transition_states() and transition_time() functions respectively]

```{r Q5}
length(unique(gapminder$year)) # we see how many unique years there are in the data set, so we can use it later on.

#Here is the code from above copy pasted with some additions.
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +#Here we make an object named anim2, where we use the illustration tool ggplot and use it on the gapminder dataset. The aestetics of the illustration is afterwards determined with gdppercap as the x-axis. lifeExp as the y-axis and population as size of the points.
  geom_point() + #We choose the style of the illustration.
  scale_x_log10() + # convert x to log scale
  transition_time(year) + # We say that the illustration should change acording to the variable year, that we find in the dataset.
  labs(title = "{closest_state}") + #Here we give illustration name(s). The name of title should be the name of the state closest to the current frame.
  transition_states(year, transition_length = 12, state_length = 1, wrap = TRUE) #Lastly we perfect when the title should change. First we decide that it should be dependent on the variable years, the the transition length (how many transitions that should be). The state_length decides how long each caption should stay up in seconds. Lastly wrap is optional and makes the illustration start over when it is finished.
anim2
```



Q6 Can you made the axes' labels and units more readable? Consider expanding the abreviated lables as well as the scientific notation in the legend and x axis to whole numbers.[hint:search disabling scientific notation]

```{r Q6}
#Here is the code from above copy pasted with some additions.
options(scipen=999) #By adding this line of code we disable the scientific way of writing lables.
anim3 <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, size = pop))+
  geom_point() + 
  scale_x_log10() + 
  transition_time(year) + 
  labs(title = "{closest_state}", x = "GDP Per Capita", y = "Life Expectancy", size = "Population") + 
  transition_states(year, transition_length = 12, state_length = 1, wrap = TRUE) +
   theme(axis.text.x = element_text(size = 12), #here we change the size of the labels.
          axis.text.y = element_text(size = 12),
          text = element_text(size = 16),
          plot.title = element_text(hjust = 0.5))
anim3
```



Q7 Come up with a question you want to answer using the gapminder data and write it down. Then, create a data visualisation that answers the question and explain how your visualization answers the question. (Example: you wish to see what was mean life expectancy across the continents in the year you were born versus your parents' birth years). [hint: if you wish to have more data than is in the filtered gapminder, you can load either the `gapminder_unfiltered` dataset and download more at https://www.gapminder.org/data/ ]


My question: Which country in Africa had the biggest drop in life expectance and how did that countries dewellopment in life expectancy look in the years from 1952-2007

```{r Q7}
africa <- ggplot(subset(gapminder, continent == "Africa"), aes( x = year, y = lifeExp, size = pop)) +
  geom_point() + 
  scale_x_log10() +
  transition_time(year) +
  transition_states(year, transition_length = 12, state_length = 1, wrap = TRUE)

africa
#Here i made a plot that looks at all the countries in Africa and their development in the years from 1952 to 2007 when we are looking at the population and its life expectancy. From the graph it is clear to see that one countries life expectancy drops dramatically. Therefore we take the next step to be sure which country that is.

buttom10 <- gapminder %>%  #We choose to look at the gapminder dataset and add pipes to it, so we can send an output of one fucntion directly to another.
  filter(year == 1992) %>%
  filter(continent == "Africa") %>% 
  select(country , lifeExp) %>% #then we select or choose to look at the countries and their gdpPercap, as everything else is irelevant right now
  arrange(desc(lifeExp)) %>%  #lastly, we arange our data in a decending order so we have the richest nation first.
  tail(10)

buttom10
#Here start by filtering the data to look at the year of the drop in life expectancy in 1992 in Africa to see which country that drops dramatically and as we see from the table it is Rwanda


rwanda <- ggplot(subset(gapminder, country == "Rwanda"), aes( x = year, y = lifeExp, size = pop, color = country)) +
  geom_point() + #geom_point() gør at vi får scatter plots altså prikker i vores kordinatsystem
  scale_x_log10() +
  transition_time(year) +
  transition_states(year, transition_length = 12, state_length = 1, wrap = TRUE)

rwanda

#As we see from the personal graph of Rwanda they have a masive drop in life expectance in 1992 but i rises quickly again. Interestingly enough the number in population does not seem to drop with the life expectance.



```